
title: "Reproducible Research: Peer Assessment 1"
author: "yanir"
output: 
  html_document:
    keep_md: true
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading and preprocessing the data
loading the data to R and libraries
```{r loaddata}
data<- read.csv("activity.csv")
library(dplyr)
library(lubridate)
library(tidyr)
library(ggplot2)
library(gridExtra)
```


## What is mean total number of steps taken per day?
preparing and summarising the data in order to make the plot
```{r makeplot}
data$date<-as_date(data$date)
agg_data<- group_by(data,date)
agg_data<- summarise(agg_data,steps_per_day=sum(steps))
plot1<- ggplot(agg_data)+geom_histogram(aes(steps_per_day),fill="white",color="blue")
plot1<-plot1+ labs(y="number of days")
plot(plot1)
```
find mean and median steps per day
```{r meanandmediansteps}
mean_steps_day<-mean(agg_data$steps_per_day,na.rm = T)
mean_steps_day
med_steps_day<- median(agg_data$steps_per_day,na.rm = T)
med_steps_day
```


## What is the average daily activity pattern?
plot average steps for each interval
```{r plotaveragesteps}
agg_data<- group_by(data,interval)
agg_data<- summarise(agg_data,avg_steps_per_interval=mean(steps,na.rm = T))
plot2<-ggplot(agg_data)+geom_line(aes(interval,avg_steps_per_interval),lwd=1.1,color="blue")
plot(plot2)
```
find interval with max steps on average
```{r findintervalwithmaxsteps}
filter(agg_data,avg_steps_per_interval==max(avg_steps_per_interval))
```

## Imputing missing values
```{r imputingNA}
#finding the NA
sapply(data,function(x){sum(is.na(x))})
##imputing avg steps where steps==NA
#calculating the avg steps for each interval
agg_data<- group_by(data,interval)
agg_data<- summarise(agg_data,avg_steps_per_interval=mean(steps,na.rm = T))
#mreging the base data with the avg steps
merge_data<- left_join(data,agg_data,by="interval")
#computing a new column to select the avg if steps=NA
new_data<- mutate(merge_data,comp_steps=ifelse(is.na(steps),avg_steps_per_interval,steps))
#see some of the new data
head(new_data)
```

plot average steps for each interval after adding avg steps insted of NA
```{r plotstepsperday}
agg_by_day<- group_by(new_data,date)
agg_by_day<-summarise(agg_by_day,comp_steps_day=sum(comp_steps))
plot3<- ggplot(agg_by_day)+geom_histogram(aes(comp_steps_day),fill="white",color="blue")
plot3<-plot3+labs(y="number of days")
plot(plot3)
```
## Are there differences in activity patterns between weekdays and weekends?
#plot average steps per interval comparing weekdays VS weekends
we will first add anew column to our data with numeric representive of the day of the week.
then we will separate  for 2 tables: one for week days and the other for weekends.
then we will plot a histogram for each table
```{r plotcomperingweekendsanweekdays}
##adding the numeric weekday
days_data<- mutate(data,week_day=wday(date))
## a table for weekdays
weekdays_data<- filter(days_data,week_day<6)
##aggragating by interval mean
agg_weekdays<- group_by(weekdays_data,interval)
agg_weekdays<- summarise(agg_weekdays,avg_steps_per_interval=mean(steps,na.rm = T))
## a table for weekends
weekdEnds_data<- filter(days_data,week_day>5)
##aggragating by interval mean
agg_weekEnds<- group_by(weekdEnds_data,interval)
agg_weekEnds<- summarise(agg_weekEnds,avg_steps_per_interval=mean(steps,na.rm = T))
plot4<-ggplot(agg_weekdays)+geom_line(aes(interval,avg_steps_per_interval),lwd=1.1,color="blue")+labs(title = "weekday")
plot4<-plot4 + scale_y_continuous(limits=c(0, 220))
plot5<-ggplot(agg_weekEnds)+geom_line(aes(interval,avg_steps_per_interval),lwd=1.1,color="blue")+labs(title = "week ends")
plot5<-plot5 + scale_y_continuous(limits=c(0, 220))
pp<-grid.arrange(plot4,plot5,nrow=2)


```

